Experience Report:

### 2010-03-16 doc/icfp-2010/main.tex
*** empty log message ***

These functions are actually derived from the generic API.

map, map\_fold, etc.

generic programming using a functor, which is then applied in each of
the 4 cases

@@

% LocalWords:  parallelize

>>>>

These five functions are actually derived from the generic API.
The generality is achieved by implementing theses derivations as a functor.
This functor, \of{Derive}, is parameterized by the \of{compute}
function, as follows:
\begin{ocaml}
module Derive
  (X : sig
     val compute : 
       f:('a -> 'b) -> 
       handle:('a * 'c -> 'b -> ('a * 'c) list) ->
       ('a * 'c) list -> unit
   end) = struct ... end
\end{ocaml}
We now explain how each function from Figure~\ref{fig:derived} is
implemented using \of{compute} inside the functor body.

\paragraph{\of{map}.} 
Function \of{map} is easily implemented by associating an integer
index with each element (of type \of{'a}) of the input list. Thus the
list of tasks passed to \of{compute} is a list of pairs of type \of{'a * int}.
As soon as an intermediate result is available, it is passed to
\of{handle}, which stores it in a local table using the index as a key
and generates no new task. Whenever \of{compute} returns, we easily
recover the output list from the values stored in the table.

\paragraph{\of{map_local_fold}.} 
Function \of{map_local_fold} is implemented using a local reference
storing the current accumulator (of type \of{'c}).
As soon as an intermediate result is available, \of{handle} 
combines it with the accumulator using function \of{fold}. Whenever \of{compute}
returns, we simply return the current accumulator.

\paragraph{\of{map_remote_fold}.}
In this case, combining the accumulator with an intermediate result is
itself a task to be performed in parallel with \of{map} operations.
Thus, when an intermediate \of{map} result is available, the
accumulator may be already in use in a \of{fold} operation. Hence we
need to store the intermediate \of{map} result in a local table, to be
used as soon as the accumulator becomes available.
The \of{handle} function distinguishes between \of{map} and \of{fold}
results and, accordingly, updates the table and the accumulator. For a
\of{map} result, either we combine it with the accumulator in a new
task or we simply store it in the local table. For a \of{fold} result,
either we simply store it into the accumulator or we immediately
combine it with a pending \of{map} result from the table.
Whenever \of{compute} returns, the accumulator must be available and
we simply return it.

\paragraph{\of{map_fold_ac}.}
Function \of{map_fold_ac} is easier than the previous one, since we do
not need to store more than one intermediate result. Indeed, as soon
as two intermediate results are available, we immediately combine them
using a \of{fold} operation. Thus the \of{handle} function 
either stores it or combines it, depending on the availability/non
availability of the accumulator.
Whenever \of{compute} returns, the accumulator must be available and
we simply return it.

\paragraph{\of{map_fold_a}.}
Function \of{map_fold_a} is definitely the trickiest. Indeed, the
operation being only associative, we can only combine \emph{adjacent}
intermediate results. In order to perform this, we store intermediate
results in a local table according to indices of the input list. More
precisely, if we have already computed
\begin{displaymath}
  x_i \oplus x_{i+1} \oplus \dots \oplus x_j
\end{displaymath}
where $\oplus$ denotes the \of{fold} operation, then we associate this
value with indices $i,j$. As soon as a result of indices $k,i-1$ or
$j+1,k$ is available, it can be combined with the result of $i,j$.
Whenever \of{compute} returns, the table should only contain a result
corresponding to the full input list and we simply return it.

@@

% LocalWords:  parallelize functor parameterized indices
